A general Nullstellensatz for generalized spaces

### 2019-03-19 paper-qcoh.tex
to do
typo
Some more nongeometric sequents

\newcommand{\xra}[1]{\xrightarrow{#1}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\bbb}{\mathfrak{b}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\B}{\mathcal{B}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\ZZ}{\mathbb{Z}}
\renewcommand{\SS}{\mathbb{S}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\op}{\mathrm{op}}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Sh}{Sh}
\DeclareMathOperator{\PSh}{PSh}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\length}{length}
\DeclareMathOperator{\List}{List}
\DeclareMathOperator{\id}{id}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Eff}{\mathrm{Ef{}f}}
\renewcommand{\_}{\mathpunct{.}\,}
\newcommand{\effective}{ef{}fective\xspace}
\newcommand{\?}{\,{:}\,}
\newcommand{\realizes}{\Vdash}
\newcommand{\notnot}{\emph{not~not}\xspace}
\usepackage{soul}\setul{0.3ex}{}
\let\oldul\ul
\renewcommand{\ul}[1]{\text{\oldul{$#1$}}}
\newcommand{\affl}{\ensuremath{{\ul{\AA}^1}}\xspace}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\brak}[1]{\llbracket #1 \rrbracket}
\newcommand{\Mod}[1]{{#1}\mathrm{\text{-}mod}}

@@

\section{Nongeometric sequents in the ring classifier}

@@

\section{On the Kock--Lawvere axiom in synthetic differential geometry}

@@

@@

* XXX Cite https://docplayer.net/55796150-Intuitionistic-algebra-some-recent-developments-in-topos-theory.html
  (in general and for internal theories)

>>>>

\newcommand{\xra}[1]{\xrightarrow{#1}}
\newcommand{\aaa}{\mathfrak{a}}
\newcommand{\bbb}{\mathfrak{b}}
\newcommand{\mmm}{\mathfrak{m}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\B}{\mathcal{B}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\OO}{\mathbb{O}}
\newcommand{\ZZ}{\mathbb{Z}}
\renewcommand{\SS}{\mathbb{S}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\defeq}{\vcentcolon=}
\newcommand{\op}{\mathrm{op}}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Sh}{Sh}
\DeclareMathOperator{\PSh}{PSh}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\length}{length}
\DeclareMathOperator{\List}{List}
\DeclareMathOperator{\id}{id}
\newcommand{\Set}{\mathrm{Set}}
\newcommand{\Eff}{\mathrm{Ef{}f}}
\renewcommand{\_}{\mathpunct{.}\,}
\newcommand{\effective}{ef{}fective\xspace}
\newcommand{\?}{\,{:}\,}
\newcommand{\realizes}{\Vdash}
\newcommand{\notnot}{\emph{not~not}\xspace}
\usepackage{soul}\setul{0.3ex}{}
\let\oldul\ul
\renewcommand{\ul}[1]{\text{\oldul{$#1$}}}
\newcommand{\affl}{\ensuremath{{\ul{\AA}^1}}\xspace}
\newcommand{\speak}[1]{\ulcorner\text{\textnormal{#1}}\urcorner}
\newcommand{\brak}[1]{\llbracket #1 \rrbracket}
\newcommand{\Mod}[1]{{#1}\mathrm{\text{-}mod}}

@@

\subsection{Nongeometric sequents in the object classifier}

Johnstone~\cite{XXX} lists several nongeometric sequents(XXX) holding in the
classifying topos of the theory~$\OO$ of objects, the theory with a single sort and
no function symbols, relations symbols or axioms, including:
\begin{align*}
  \forall x,y\?U_\OO\_ \neg\neg(x = y) &&
  \forall x_1,\ldots,x_n\?U_\OO\_ \neg \forall y\?U_\OO\_ \bigvee_{i=1}^n (y = x_i)
\end{align*}
Intuitively, the first expresses that~$U_\OO$ is close to a subsingleton, while
the second expresses that~$U_\OO$ is close to being infinite. Using the
Nullstellensatz, these formulas can be verified as follows.

We argue internally to~$\Set[\OO]$. Let~$x,y\?U_\OO$ and assume~$\neg(x = y)$.
Then~$\ul{\OO}/U_\OO$ proves~$(x = y \seq{[]} \bot)$. Hence this sequent holds
of any model of~$\ul{\OO}/U_\OO$, in particular of the
quotient~$U_\OO/(x \sim y)$. Hence~$\bot$ holds for this model, that is, we
have a contradiction.

For the second formula, let~$x_1,\ldots,x_n\?U_\OO$ and assume~$\forall
y\?U_\OO\_ \bigvee_{i=1}^n (y = x_i)$. By the Nullstellensatz, this universal
statement is provable and hence holds for every~$\ul{\OO}/U_\OO$-model. But it
does not hold for~$U_\OO \amalg \{\star\}$.

The argument using the Nullstellensatz makes it transparent that both formulas
are also satisfied by the generic model of any Horn theory~$\TT$: In that case,
instead of constructing the quotient set~$U_\OO/(x \sim y)$, we construct the
quotient model~$(\ul{\TT}/U_\TT)\langle \,|\, x = y \rangle$; and instead of~$U_\OO
\amalg \{\star\}$, we construct the free model~$(\ul{\TT}/U_\TT)\langle \star
\,|\, \rangle$ (XXX need to exclude trivial singleton theory).

\subsection{Nongeometric sequents in the ring classifier}

@@

\subsection{On the Kock--Lawvere axiom in synthetic differential geometry}

@@

% XXX object classifier and related

@@
